---
title: "Risk-Parity Portfolio"
output: html_document
number_section: TRUE
---
***
# Inrtoduction
Practical flaws of Markowitz Model:  

1.variance is not a good risk measure in practice. The solution is use alternative measure for risk, such as downside risk (VaR, CVaR, Sortino ratio). The theoreticians described a number of properties a good risk measure should satify: __a coherent risk measure__. Such properties include monotonicity, sub-additivity, homogeneity, and translational invariance.
_Ref_ [coherent risk measure](https://en.wikipedia.org/wiki/Coherent_risk_measure)

2.Sensitivity w.r.t estimators (parameter estimation of covariance matrix and mean vector by using historic data). If we use historic data, we introduce another assumption: no time-varying of estimator(and in practice Smaple size cannot be large enough due to lack of stationarity of data). Some improvements includes modeling \(r_t\) conditional on
\(\mathcal{F_{t-1}}\) and give different weights to observation data like __Garch__ model by not assuming that conditional covariance and conditional mean are constant.
Solutions include __shrinkage estimators__, __Black-Litterman estimators__ (similar to a reverse engineering: do things backwards).  

3.Lack of diversification (this seems quite contradictive; but in most case we allocate a large weight to only few assets): concentrates risk too much in few assets (2008 financial crisis): solution is the __risk-parity portfolio__.

### Conclusion
Simply speaking, Markowitz framework is two-objectives optimiation problem: the mean return \(\mathbf{w^T}{\boldsymbol{\mu}}\) and the variance \(\mathbf{w^T}\mathbf{\Sigma}\mathbf{w^T}\). The idea is to find a trade-off between these two regarding how risk-averse a investor is.

__Risk parity portfolio__  

* Less senstive to parameter estimation errors
* Diversifies the risk


# Review: Emprical demonstration of flaws of Markowiz framwork



\[\begin{array}{} \underset{\mathbf{w}}{maximize} & \boldsymbol{\mu}^T - \lambda {\mathbf{w}^T}{\mathbf{\Sigma w}}\\
{subject~ to} &\mathbf{1}^T \mathbf{w} = 1\\
&\mathbf{w}\geqslant 0 \end{array}\]


For convience, we define a function to find the solution using _CVXR_  package for convex quadratic optimization problem.

```{r weights_Markowitz, message=FALSE, warning=FALSE}
library("CVXR")
weights_Markowitz <- function(lambda, Sigma, mu){

  w <- Variable(length(mu))
  res <- solve(Problem(Maximize(t(mu) %*% w - lambda*quad_form(w, Sigma)),
                       constraints = list(w >= 0, sum(w) == 1)))
  w_Markowitz <- as.matrix(res$getValue(w))
  rownames(w_Markowitz) <- stock_namels
  return(w_Markowitz)
}
```

```{r Data collection, message=FALSE, warning=FALSE}
library(xts)
library(quantmod)
begin_date <- "2015-01-01"
end_date <- "2019-07-13"
stock_namels <- c("AAPL", "IBM", "AMZN", "BAC", "MS", "FDX", "MMM", "KO", "PFE")
prices <- xts()
stock = "IBM"
for (stock in stock_namels){
  prices <- cbind(prices, Ad(getSymbols(stock, from = begin_date, to = end_date, auto.assign = FALSE)))
}

plot(prices/rep(prices[1, ], each = nrow(prices)), legend.loc = "topleft",
     main = "Normalized prices")
logret <- diff(log(prices))
colnames(logret) <- stock_namels
mu <- colMeans(logret, na.rm = TRUE)
Sigma <- cov(na.omit(logret))
```
## Lack of diversification
Generally, \(\lambda\) is between 0 and 4. The larger \(\lambda\), the more risk-averse.

```{r lack_diversification }
par(mfrow = c(3,2))

for (lmd in c(0.5,1,2,4,8,10)){
  barplot(t(weights_Markowitz(lmd, Sigma, mu)), main = "Portfolio allocation", xlab = paste("lambda = ", lmd))
}
par(mfrow = c(1,1))

```
__Remark__: if \(\lambda\) goes to positive infinity, it will converge to GMVP(Global Minimum Variance Portfolio) situation


#Risk-Parity Portfolio -- most simple case 

_More complicated case might be added if I study further_

\[\sigma (\mathbf{w}) = \sqrt{\mathbf{w^T}\mathbf{\Sigma}\mathbf{w}}\]

\[= \sum_{i=1}^{N} w_i \frac{\partial\sigma}{\partial w_i}\] 
\[ =  \sum_{i=1}^{N}\frac{w_i\mathbf{(\Sigma w )}_i}{\sqrt{\mathbf{w^T}\mathbf{\Sigma}\mathbf{w}}}  \]

__Normalized risk contribution__: \(\frac{w_i\mathbf{(\Sigma w )}_i}{\sqrt{\mathbf{w^T}\mathbf{\Sigma}\mathbf{w}}}  \)  

__risk budgeting equations__
\[w_i \mathbf{(\Sigma w)_i} = b_i \mathbf{w^T\Sigma w}, \qquad  i = 1,...,N \]

__Risk parity portfolio__ is a special case given the budgeting portfolio with  \(b = 1/N\), more specificly, the risk budgeting portfolio (no short position)is:
\[w_i = \frac {\sqrt{b_i}/\sqrt{\Sigma_{ii}}}{\sum_{k=1}^{N}\sqrt{b_k}/\sqrt{\Sigma_{kk}}}, \qquad i =1, ... , N.  \]



## Closed-form solution for most simple case

```{r Closed-form solution}
w_naive <- 1/sqrt(diag(Sigma))
w_naive <- w_naive/sum(w_naive)
w_Markowitz_lmd1 <- weights_Markowitz(1 , Sigma, mu)
w_Markowitz_lmd10 <- weights_Markowitz(10 , Sigma, mu) 
w_all <-  cbind(w_Markowitz_lmd1,w_Markowitz_lmd10, w_naive)
colnames(w_all) <- c("Mkwtz_lmd1", "Mkwtz_lmd10","risk-parity" )
barplot(t(w_all), beside = TRUE, legend = colnames(w_all))

```

# Test for outsample performance (waited to be added)

# Sensitivity to parameters  (particularly to \(\mu\))(waited to be added)




